---
title: "Homework 5"
author: "Riyadh Baksh"
date: "`r Sys.Date()`"
output: github_document
---

## Problem 1

The plot below shows that in general, the probability of having two people with the same birthday increases as the group size increases. The probability increase rate is faster as the group size increases from small sizes, but flattens out as the group size gets larger.

```{r, message=FALSE}

library(tidyverse)
library(rvest)

set.seed(1)

bday = function(n) {
  days = sample(1:365,size=n,replace=TRUE)
  decision=FALSE
  for (i in 1:n) {
    for (j in 1:n) {
      if (i!=j) {
        if (days[i]==days[j]){
          decision=TRUE
        }
      }
    }
  }
  decision
}

prob = vector("numeric",length=49)

for (n in 2:50) {
  output = vector("numeric",length=10000)
  for (i in 1:10000) {
    output[i] = bday(n)
  }
  prob[n-1] = mean(output)
}

ggplot(mapping=aes(x=2:50,y=prob)) +
  geom_point() +
  xlab("Group Size") +
  ylab("Probability of Duplicate Birthdays")
```

## Problem 2

The code is shown below for this problem.

```{r}

normal = function(u) {
  result=tibble(estimate=numeric(),p.value=numeric(),true=numeric())
  for (i in 1:5000){
    x = rnorm(30,mean=u,sd=5)
    test = broom::tidy(t.test(x))
    result = add_row(result,
                     estimate=as.numeric(test[1]),
                     p.value=as.numeric(test[3]),
                     true=u)
  }
  result
}

values = tibble(estimate=numeric(),p.value=numeric(),true=numeric())

for (u in 0:6){
  result=normal(u)
  values = add_row(values, result)
}

```

The graph below shows the power of the test compared to the true value of $\mu$. The null hypothesis is fixed as $\mu=0$, and therefore, as the true value of $\mu$ increases from 1, the power quickly increases and then starts to plateau until it reaches essentially 100% at $\mu=6$.

```{r, message=FALSE}

data =
  values |>
  mutate(test=p.value<=0.05)

data |>
  group_by(true) |>
  summarize(
    power = mean(test)
  ) |>
  ggplot(aes(x=true,y=power)) +
    geom_point() +
    geom_smooth(se=FALSE) +
    xlab("True Value of μ") +
    ylab("Power") +
    ggtitle("Null Hypothesis: μ=0")
```

As the first plot below shows, when the average estimates of $\mu$ are plotted against the true value, they essentially equal each other, since the points fall along the line $y=x$. This makes sense because many samples are taken from the distribution with the true $\mu$.

However, in the second plot, when we only consider samples that reject the null hypothesis, the average estimate needs to be greater than the true value at values close to the null. This provides sufficient evidence to reject the null. That is, when the true value is $\mu=1$ for example, samples generated will likely overlap with zero, making it harder to reject the null. However, at values farther from the null, the null is typically rejected anyways, so the average estimate approximates the true value (reflects the first plot since many samples are not filtered out). The one difference is that when the true value is zero, there aren't many samples that reject the null, so the average of the samples that *do* reject would still have to be higher than zero, but not that much since the distribution which these samples came from is still centered around zero.

```{r, message=FALSE}

library(patchwork)

first =
  data |>
  group_by(true) |>
  summarize(
    average = mean(estimate)
  ) |>
  ggplot(aes(x=true,y=average)) +
    geom_point() +
    geom_smooth(se=FALSE) +
    geom_line(aes(x=0:6,y=0:6)) +
    xlab("True Value of μ") +
    ylab("Average Estimate of μ") +
    ggtitle("All samples")

second =
  data |>
  filter(test==TRUE) |>
  group_by(true) |>
  summarize(
    average = mean(estimate)
  ) |>
  ggplot(aes(x=true,y=average)) +
    geom_point() +
    geom_smooth(se=FALSE) +
    geom_line(aes(x=0:6,y=0:6)) +
    xlab("True Value of μ") +
    ylab("Average Estimate of μ") +
    ggtitle("Only samples rejecting null of μ=0")

first + second


```

## Problem 3

The table below shows the number of solved and unsolved homicides across all cities.

```{r, message=FALSE}

data =
  read_csv(file="data/homicide.csv") |>
  mutate(
    city_state = paste(city,state),
    type = case_match(
      disposition,
      "Open/No arrest" ~ "Unsolved",
      "Closed by arrest" ~ "Solved",
      "Closed without arrest" ~ "Unsolved",
    )
  ) |>
  select(city_state,type)

data |>
  group_by(type) |>
  summarize(homicides = n()) |>
  knitr::kable()

```

The table below shows the estimate and 95% confidence interval for the proportion of unsolved homicides in Baltimore, MD.

```{r}

baltimore =
  data |>
  filter(city_state=="Baltimore MD")

n_unsolved =
  baltimore |>
  select(type) |>
  count(type,name="count") |>
  filter(type=="Unsolved") |>
  select(count) |>
  as.numeric()

n_total =
  baltimore |>
  count(city_state,name="count") |>
  select(count) |>
  as.numeric()
  

output = broom::tidy(prop.test(x=n_unsolved,n=n_total))

output |>
  select(estimate,conf.low,conf.high) |>
  knitr::kable()

```

The graph below shows the estimate for the proportion of unsolved homicides in each city along with their 95% confidence interval. The graph is sorted from greatest proportion to least.

```{r}

prop = function(citystate){

city =
  data |>
  filter(city_state==citystate)

n_unsolved =
  city |>
  select(type) |>
  count(type,name="count") |>
  filter(type=="Unsolved") |>
  select(count) |>
  as.numeric()

n_total =
  city |>
  count(city_state,name="count") |>
  select(count) |>
  as.numeric()

output = broom::tidy(prop.test(x=n_unsolved,n=n_total))

output |>
  select(estimate,conf.low,conf.high)
}

dist_cities =
  data |>
  distinct(city_state) |>
  select(city_state) |>
  filter(city_state!="Tulsa AL") # There seems to be a mistake in coding because there is no Tulsa in Alabama, and even if, there isn't enough data for a proportion test for "Tulsa AL".

output = map_dfr(dist_cities$city_state,prop)

output =
  output |>
  mutate(dist_cities)

ggplot(output,aes(x=reorder(city_state,-estimate),y=estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low,ymax=conf.high)) +
  theme(axis.text.x=element_text(angle=90)) +
  xlab("City") +
  ylab("Proportion of Unsolved Homicides")
```
